---
title: "My Computational Musicology Dashboard"
author: "saryanasser"
date: "Block 4 - 2025"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: cosmo
---
```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(plotly)
```

### Introduction

Welcome to my Computational Musicology portfolio for 2025! This storyboard contains further examples from each week to inspire you.

### Visualising the AI Song Contest

```{r}
aisc2024 <- read_csv("/Users/saryanasser/Library/CloudStorage/OneDrive-UvA/Computational Musicology/saryanasser/aisc2024.csv")
aisc2024 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = tempo,
      y = arousal,
      size = instrumentalness,
      colour = danceability
    )
  ) +
  geom_point() +              # Scatter plot.
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  geom_text(                  # Add text labels from above.
    x = 121,
    y = 4.91,
    label = "Onda Corta - Sud America",
    size = 3,                 # Override size (not loudness here).
    hjust = "left",           # Align left side of label with the point.
    vjust = "center",         # Align vertical centre of label with the point.
    angle = 30                # Rotate the text label
  ) +
  scale_x_continuous(         # Fine-tune the x axis.
    limits = c(50, 200),
    breaks = c(50, 100, 150, 200), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(1, 9),
    breaks = c(1, 5, 9),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Tempo",
    y = "Arousal",
    colour = "Danceability"
  )
ggplotly()
```
***
This is the bad visualisation of the AI Song Contest we used in our first lab session, this time in a dashboard.

### Visualising the AI Song Contest 2
```{r}
aisc2024 |> 
  ggplot(aes(x = tempo, y = arousal, color = danceability)) +
  geom_point(alpha = 0.7, aes(size = instrumentalness)) +  # Add transparency for visibility
  geom_smooth(method = "loess", color = "black", linetype = "dashed", se = FALSE) +  # Trend line for clarity
  scale_x_continuous(
    limits = c(50, 200),
    breaks = seq(50, 200, by = 50)
  ) +
  scale_y_continuous(
    limits = c(1, 9),
    breaks = c(1, 5, 9)
  ) +
  scale_color_viridis_c(option = "plasma") +  # Use a distinct, high-contrast palette
  scale_size(range = c(1, 6)) +  # Keep point sizes meaningful
  theme_minimal(base_size = 14) +  # Cleaner theme with larger text
  labs(
    title = "AI Song Contest 2024: Tempo vs. Arousal",
    subtitle = "Danceability indicated by color, instrumentalness by size",
    x = "Tempo (BPM)",
    y = "Arousal (Energy Level)",
    color = "Danceability",
    size = "Instrumentalness"
  ) +
  theme(legend.position = "bottom")  # Move legend to avoid overlap
```

***
New and updated version 

### Personal track descriptions

I decided on exploring different genres that I like by asking different models for various genres of songs. I have always been a fan of post-punk, alternative rock music from the late 90s to early 2000s, such as Interpol, The Strokes, Bloc Party, Arctic Monkeys, Fontaines D.C. With the help of AI, I came up with this description to use as a prompt for gen AI music models: a high-energy alt rock / post-punk song with a melodic bassline, intricate drumming, and sharp and rhythmic guitar work, reminiscent of the bands Interpol and Bloc Party. dynamic, with tension-building verses leading into an explosive, anthemic chorus. create a sense of depth and intensity.” I also used this shortened version for models with character limits “Post-Punk, Driving Melodic Bassline, Angular Reverb-Drenched Guitars, Punchy Dynamic Drumming, Moody Detached Vocals, Urgent & Anthemic, Dark Yet Energetic, Tension-Building Composition, 140 BPM”. I have also recently been enjoying deep house music, so I decided to choose this genre as one of my songs. I used the following prompts, “deep house song that has a hypnotic beat, gradually layering warm synths, deep basslines, and subtle percussive beats, with a steady, entrancing rhythm. slow, cinematic build-ups that evoke nostalgia and euphoria. Incorporate atmospheric pads and a shimmering, time-dissolving feel of the track, with immersive, and emotionally uplifting verses and bridges, suitable for a sunset in the mountains” and “Deep House, Hypnotic Synth Pads, Pulsing Bassline, Rolling Four-on-the-Floor Groove, Atmospheric Textures, Slow-Building Progression, Dreamy Vocal Samples, Cinematic, Nostalgic, Expansive, 120 BPM”. I wanted to try a different genre that I also enjoy, something along the lines of Lana del Rey’s style. So I used the following prompts, “A cinematic baroque / dream pop composition that blends dreamy electronic synths with classical orchestration. Feature violins, melancholic clarinets, and rich trumpet swells, weaving through ethereal synth pads and delicate, reverberated piano. The rhythm should be slow and hypnotic, with a hazy, dreamlike quality. The vocals should be intimate yet grand, drenched in vintage-style reverb, with poetic, melancholic lyrics evoking themes of romance, nostalgia, and faded Hollywood glamour. Think of Lana Del Rey’s storytelling style, but with a modern dream pop twist—layered harmonies, sweeping crescendos, and an air of cinematic longing .” and “Baroque Pop / Dream Pop, Ethereal Synth Pads, Sweeping Violin & Clarinet Arrangements, Melancholic Trumpet Swells, Reverb-Drenched Intimate Vocals, Vintage Aesthetic, Poetic & Nostalgic, Cinematic & Grand, 80 BPM”. I explored the outputs of these prompts from various models including Suno, Stable Audio, Beatoven.ai, Soundverse.ai, Udio, and Mubert. Although I was hesitant to try Suno and Udio given their use of artists’ music without compensating them, I wanted to see whether there would be any differences in the quality, production output, relevance to the prompt, and similarity to expectations and existing songs. I ended up deciding on the Stable Audio deep house track because it seemed to best match my expectations of emotive, intense, while also calming and not sounding too elaborate. I found the vocal and lyrical qualities of most models to be of somewhat lower quality than I was expecting, with many songs sounding unnatural or AI generated (understandably). I ended up deciding to go with another deep house song that I generated on Suno.  


### Conclusion / Discussion

n/a